Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis

### 2017-12-07 paper.tex
Added a few equations and some text to Budget Computation section.

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   N_m^2 = \frac{g}{T}(\frac{\partial T}{\partial z}+\Gamma_m)(1+\frac{T}{R_d/R_v+q_s}\frac{\partial q_s}{\partial T})-\frac{g}{1+q_t}\frac{\partial q_t}{\partial z},
   \end{equation}
where g is gravitational acceleration, T is temperature, $R_d$ and $R_v$ are the gas constants of dry air and water vapor, respectively, $q_s$ is the saturation mixing ratio, and $q_t$ is the total condensate mixing ratio. $\Gamma_m$ is the moist-adiabatic lapse rate:
   \begin{equation} \label{eq:gamma_m}
   \Gamma_m = DETERMINE WHAT THIS IS FOR THE NON-CONSERVATIVE EQUATION SET
   \end {equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ approach zero. In this limiting case, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where \textit{g} is gravitational acceleration and $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in unsaturated environments; for mathematical simplicity, however, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in this section.}.

Taking the time derivative of Eq. \ref{eq:n2dry} yields the static stability tendency:
   \begin{equation} \label{eq:dn2dt}
   \frac{\partial N^2}{\partial t} = \frac{g}{\theta}\frac{\partial}{\partial z}\frac{\partial \theta}{\partial t}+\frac{g}{\theta^2}\frac{\partial \theta}{\partial z}\frac{\partial \theta}{\partial t},
   \end{equation}
where the potential temperature tendency, $\frac{\partial \theta}{\partial z}$, is:
   \begin{equation} \label{eq:dthetadt}
   \frac{\partial \theta}{\partial t} = 
   \end{equation}
Each term on the right-hand side of Eq. \ref{eq:dthetadt} represents a potential temperature budget variable, each of which is output directly by the model every minute.
These tendencies are substituted for $\frac{\partial \theta}{\partial t}$ in Eq. \ref{eq:dn2dt}, yielding the static stability tendency attributable to each budget term every minute.
 \section{Results}

>>>>

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   N_m^2 = \frac{g}{T}(\frac{\partial T}{\partial z}+\Gamma_m)(1+\frac{T}{R_d/R_v+q_s}\frac{\partial q_s}{\partial T})-\frac{g}{1+q_t}\frac{\partial q_t}{\partial z},
   \end{equation}
where $g$ is gravitational acceleration, $T$ is temperature, $R_d$ and $R_v$ are the gas constants of dry air and water vapor, respectively, $q_s$ is the saturation mixing ratio, $q_t$ is the total condensate mixing ratio, and $\Gamma_m$ is the moist-adiabatic lapse rate:
   \begin{equation} \label{eq:gamma_m}
   \Gamma_m = DETERMINE WHAT THIS IS FOR THE NON-CONSERVATIVE EQUATION SET
   \end {equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ approach zero. In this limiting case, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in sub-saturated environments; for mathematical simplicity, however, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in this section.}.

Taking the time derivative of Eq. \ref{eq:n2dry} yields the static stability tendency:
   \begin{equation} \label{eq:dn2dt}
   \frac{\partial N^2}{\partial t} = \frac{g}{\theta}\frac{\partial}{\partial z}\frac{\partial \theta}{\partial t}+\frac{g}{\theta^2}\frac{\partial \theta}{\partial z}\frac{\partial \theta}{\partial t},
   \end{equation}
where the potential temperature tendency, $\frac{\partial \theta}{\partial z}$, is:
   \begin{equation} \label{eq:dthetadt}
   \frac{\partial \theta}{\partial t} = HADV+VADV+HTURB+VTURB+MP+RAD+DISS 
   \end{equation}
Each term on the right-hand side of Eq. \ref{eq:dthetadt} represents a potential temperature budget variable, each of which is output directly by the model every minute.
HADV and VADV are the radial and vertical advective tendencies, HTURB and VTURB are the radial and vertical tendencies from the turbulence parameterization, MP is the tendency from the microphysics scheme, RAD is the tendency from the radiation scheme, and DISS is the tendency due to turbulent dissipation.
This equation neglects Rayleigh damping, since this term is zero everywhere below 25 km, and the analysis domain does not extend to that height.
Each term in Eq. \ref{eq:dthetadt} is substituted for $\frac{\partial \theta}{\partial t}$ in Eq. \ref{eq:dn2dt}, yielding the contribution of each budget term to the static stability tendency.
These terms are then summed, yielding an instantaneous "budget change" in $N^2$ every minute.
These instantaneous budget changes are then averaged over 24-hour periods and compared to the total model change in $N^2$ over that same time period, i.e.:
   \begin{equation} \label{eq:budgetchange}
   \Delta N^2_{budget} = \sum_{t=t_0}^{t_0+\delta t} \frac{\partial N^2}{\partial t}\mid_t, NEED BAR OVER SUMMATION
   \end{equation}
   \begin{equation} \label{eq:modelchange}
   \Delta N^2_{model} = N^2_{t_0+\delta t}-N^2_{t_0},
   \end{equation}
where $t_0$ is an initial time and $\delta_t$ is 24 hours.
In Eq. \ref{eq:budgetchange}, Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in sub-saturated environments; $t_0$ is an initial time and $\delta_t$ is 24 hours.

 \section{Results}
